---
title: "NBA_Final_Project"
author: "Ye Chao"
date: "2019年6月3日"
output: html_document
---
## Library:
```{r}
#Library:
library("data.table")
library("dplyr")
library("mgcv")
library("MASS")
library("corrplot")
library("RColorBrewer")
library("caret")
library("Metrics")
library("glmnet")
library("lars")
library("ggplot2")
library("usmap")
library("tidyverse")
library("pls")
library("psych")
library("SensoMineR")
```

##Importing the data:
```{r}
##======================================================================================##
##Importing the data##
Seasondata <- read.csv("C:/Users/chaoye/Desktop/Davis/Quarter 3/TTP 289/project/data/Seasons_Stats.csv", header = TRUE)
#str(Seasondata)
Seasondata_Complete <- Seasondata[complete.cases(Seasondata),]
Seasondata_After1980 <- subset(Seasondata, Seasondata$Year >= 1980)

Players <- read.csv("C:/Users/chaoye/Desktop/Davis/Quarter 3/TTP 289/project/data/Players.csv", header = TRUE)
Salary <- read.csv("C:/Users/chaoye/Desktop/Davis/Quarter 3/TTP 289/project/data/salary.csv", header = TRUE)
Seasondata_full <- merge(Seasondata_Complete, Players, by = "Player",all.x = TRUE)
#Seasondata_full <- arrange(b, Year)
Seasondata_full$Average.points <- Seasondata_full$Points/Seasondata_full$Games


a <- subset(Seasondata_full, Seasondata_full$Year >= 1990)

Season_salary <- merge(a, Salary, by = c("Player", "Year"),all = TRUE)
Season_salary <- Season_salary[complete.cases(Season_salary),]
Season_salary=Season_salary[!duplicated(Season_salary),]
Season_salary$Salary.in.. = as.numeric(gsub("[\\$,]", "", Season_salary$Salary.in..))

rm(a)
##======================================================================================##
```


## mtable Function:
```{r}
mtable<- function (x, bycol, firstvarcol){
#A function to compute a means table for a matrix.
#x - the data frame with the data
#bycol - the row or rows used for grouping (usually products)
#use c(col1,col2) as the bycol option if using more than one column.
#firstvarcol - the col containing the first variable
if (length(bycol)==1){
  mns<-matrix(nrow=0, ncol=length(levels(as.factor(x[,bycol]))))
  for (n in firstvarcol:length(x)){
    m.r<-with(x, tapply(x[,n], x[,bycol], mean))
    mns<-rbind(mns,m.r[])
    }
  mns<-as.data.frame(mns)
  names(mns)<-names(m.r)
  rownames(mns)<-names(x[firstvarcol:length(x)])
  mns<-t(mns)
  return(mns)
  }
else{
  bc<-paste(x[,bycol[1]],names(x)[bycol[2]],x[,bycol[2]])
  x.2<-as.data.frame(cbind(bc,x))
  mns<-matrix(nrow=0, ncol=length(levels(as.factor(x.2$bc))))
  
  for (n in (firstvarcol+1):length(x.2)){
    m.r<-with(x.2, tapply(x.2[,n], x.2$bc, mean))
    mns<-rbind(mns,m.r[])
  }
  mns<-as.data.frame(mns)
  names(mns)<-names(m.r)
  rownames(mns)<-names(x.2[(firstvarcol+1):length(x.2)])
  mns<-t(mns)
  titl<-paste("Means by", names(x)[bycol[1]], "and",
  names(x)[bycol[2]])
  return(mns)
  }
}
```


## Part 1: General Analysis:

#### Average Points by Year:
```{r}
Seasondata_After1980$Averagepoints <- (Seasondata_After1980$Points/Seasondata_After1980$Games)

P1 <- ggplot(Seasondata_After1980, aes(x= as.factor(Year),y=Averagepoints)) +
  geom_boxplot()+
  labs(x= "Season",y = "All Players Average Points ")+
  labs(title = 'Average points all players', subtitle = "by year since 1980")

P1
```

#### Score Way by Year
```{r}
a <- tapply(Seasondata_After1980$X3point.fieldgoal.attempts, (as.factor(Seasondata_After1980$Year)), mean)
A <- data.frame(c(1980:2017),a,"3points.shooting")
names(A) <- c("year","times", "way")

b <- tapply(Seasondata_After1980$X2point.fieldgoal.attempt, (as.factor(Seasondata_After1980$Year)), mean)
B <- data.frame(c(1980:2017),b,"2points.shooting")
names(B) <- c("year","times", "way")

c <- tapply(Seasondata_After1980$Free.throw.attempts, as.factor(Seasondata_After1980$Year), mean)
C <- data.frame(c(1980:2017),c,"freethrow")
names(C) <- c("year","times", "way")

p2 <- rbind(rbind(A,B),C)

P2 <- ggplot(p2, aes(x=year,y=times))+
  geom_bar(aes(fill=p2$way),position="dodge",stat = "identity" )+
  labs(title = "Way of scores", subtitle = "1980-2017")+
  labs(fill = "Way of Scores")
P2

rm(a)
rm(b)
rm(c)
rm(A)
rm(B)
rm(C)
rm(p2)

```

#### University Data
```{r}
#### University sta
a<- as.data.frame(table(na.omit(Season_salary$collage)))
a <- a[order(a$Freq),]
# nrow(a)
a <- a[424:1,]
a <-a[-1,]
names(a) <- c("college","Freq")
a <-a[1:10,]

a$college=factor(a$college, levels=c("Syracuse University","Michigan State University","University of Michigan","University of Kansas","Georgia Institute of Technology","Duke University","University of Arizona","University of California, Los Angeles","University of Kentucky","University of North Carolina"))

P3 <- ggplot(a, aes(x = college , y=Freq))+
  geom_col()+
  ylab("Number of Students Played in NBA")+
  coord_flip()+
  theme(text = element_text(size=20),
              axis.text.x = element_text(angle=60, hjust=1)) +
labs(title = "Top 10 colleges whose students get to play in NBA since 1980")
P3
rm(a)

```

## Part 2: Linear Models

#### A Failed model:
```{r}

model1 <- lm(Win.share48 ~  height + weight + Age + Minutes.played + Total.rebounds + Free.throw.attempts + X2point.fieldgoal.attempt + X3point.fieldgoal.attempts + Effective.fieldgoal.percetage + Personal.fouls + Steal + Blocks,  data = Seasondata_full)
summary(model1)

# Win share.48 is likely calculated by those variables, so the salary data was brought in. 

```

#### Cor-relation Plot:
```{r}
x <- data.frame(Seasondata_full[,-(1:4)])
x <- x[,-c(2,4,18,19,20,51,52,53,54)]
x<- x[,-c(42,43)]
x<- as.data.frame(x)

corrplot(cor(x), type="lower", order="hclust", col=brewer.pal(n=8, name="RdYlBu"), tl.cex = 0.7, tl.col = "black")
rm(x)
```

#### Define "real.salary":
```{r}
# annual 2% currency inflation rate is considered.
Season_salary$real.salary <- Season_salary$Salary.in..* 1.02^ ( 2018- (Season_salary$Year))
real.salary <- Season_salary$real.salary
```


#### Principle Component Analysis all Variables:

```{r}
# PCA----all variables

######.full includes "real.salary", Capital means data.matrix ######

pca.x <- data.frame(Season_salary$Age,Season_salary$Games, Season_salary$True.shooting.percetage, Season_salary$Offensive.rebound.percetage, Season_salary$Defensive.rebound.percetage, Season_salary$Assist.percetage, Season_salary$Steal.percetage, Season_salary$Block.percetage, Season_salary$Turnover.percetage, Season_salary$Field.goal.percetage, Season_salary$X3point.fieldgoal.percetage, Season_salary$X2point.fieldgoal.percetage, Season_salary$Free.throw.percetage, Season_salary$Average.points, Season_salary$height, Season_salary$weight)

pca.full <- data.frame(real.salary,pca.x)
# 
pca.X <- data.matrix(pca.x)
pca.FULL <- data.matrix(pca.full)

fa.parallel(pca.x, fa = 'pc', n.iter = 100, show.legend = FALSE) #Scree plot: outputing number of components

pca = PCA(pca.x, scale.unit=TRUE, ncp=4, graph=TRUE) # Variables Factor Map(PCA), first quadrant: 2point field goal percetage, true shooting percetage, field goal percetage, season salary age, 3 point field goal percetage, games, field goad percetage, season salary.      
### (Multiple collinearity)####

#pca = PCA(pca.x, scale.unit=FALSE, ncp=4, graph=TRUE)

#dimdesc(pca, axes=c(1,2))

da.panelellipses=panellipse(pca.x, col.p=2, col.j=1, firstvar=3) # Variables Factor Map(PCA), first quadrant: 2point field goal percetage, true shooting percetage, field goal percetage, season salary age, 3 point field goal percetage, games, field goad percetage, season salary.      
### (Multiple collinearity)####

# pca.method1 <- principal(pca.x, nfactors = 4, rotate = "none")
# pca.method1
# 
# pca.model <- lm(real.salary ~ pca.method1$scores[,1] + pca.method1$scores[,2])
# summary(model1)
# 
# pca.cv <- data.frame(real.salary, pca.method1$scores[,1], pca.method1$scores[,2])
# names(pca.cv) <- c("real.salary", "pc1", "pc2")
# 
# pca.result = train(real.salary ~ pc1 + pc2, data= pca.cv, trControl = k10_cv, method = "lm")
# pca.result$results
```

#### lm salary try
```{r}
##### Model 2: Salary: lm

model2<- lm(formula = real.salary ~ Age  + Total.rebound.percetage + 
    Assist.percetage + Steal.percetage + Field.goal.percetage + Free.throw.percetage + height + Average.points, 
data = Season_salary)

summary(model2)
plot(model2)

# leverage1 <- hat(model.matrix(model2))
# plot(leverage1, pch = 19, ylab = "leverage", main = "lm1 leverage plot with odo faulty colored red")
# 
# cook <- cooks.distance(model2)
# plot(cook, pch = 19, ylab = "cook's distance", main = "lm1 cook's distance plot with odo faulty colored red")


x <- data.frame(Season_salary$Age, Season_salary$Player.efficiency.rating,Season_salary$Total.rebound.percetage ,
    Season_salary$Assist.percetage , Season_salary$Steal.percetage , Season_salary$Field.goal.percetage , 
    Season_salary$Free.throw.percetage , Season_salary$height)
X <- data.matrix(x)



set.seed(123)
training.samples <- real.salary %>%
  createDataPartition(p = 0.7, list = FALSE)
train.data  <- x[training.samples, ]
train.y <- (real.salary[training.samples])
test.y <- real.salary[-training.samples]
test.data <- x[-training.samples, ]
# Build the model
model2 <- lm(real.salary ~ X, data = train.data)
# Make predictions and compute the R2, RMSE and MAE
predictions <- model2 %>% predict(test.data)

#MSE:
#data.frame( R2 = R2(predictions, real.salary),
model2_RMSE = RMSE(predictions, real.salary)
model2_MSE = mse(predictions, real.salary)



k10_cv = trainControl(method="cv", number=10)
ols_model2 = train((real.salary) ~ Age  + Total.rebound.percetage + 
    Assist.percetage + Steal.percetage + Field.goal.percetage + 
    Free.throw.percetage + height  + Average.points, 
    data = Season_salary,trControl = k10_cv, method = "lm")
ols_model2$results

result1 <- as.data.frame(ols_model2$results)


```

## Try something different to solve the potential Multiple collinearity Problem
```{r}
##### Model 2: Salary: lm
model3 <- lm (real.salary ~  weight + True.shooting.percetage + Turnover.percetage + Blocks, data = Season_salary)
summary(model3)


stepAIC(model3, direction = "backward")


model3<- lm(formula = real.salary ~ True.shooting.percetage + Turnover.percetage + Blocks, data = Season_salary)
summary(model3)
plot(model3)


x <- data.frame(Season_salary$True.shooting.percetage, Season_salary$Turnover.percetage, Season_salary$Turnover.percetage, Season_salary$Blocks)
X <- data.matrix(x)



set.seed(123)
training.samples <- real.salary %>%
  createDataPartition(p = 0.7, list = FALSE)
train.data  <- x[training.samples, ]
test.data <- x[-training.samples, ]
# Build the model
model3 <- lm(real.salary ~ X, data = train.data)
# Make predictions and compute the R2, RMSE and MAE
predictions <- model3 %>% predict(test.data)

#MSE:
RMSE_model3 = RMSE(predictions, real.salary)
MSE_model3 = mse(predictions, real.salary)

k10_cv = trainControl(method="cv", number=10)
ols_model3 = train(real.salary ~ True.shooting.percetage + Turnover.percetage + Blocks, data = Season_salary,trControl = k10_cv, method = "lm")
ols_model3$results

# But it doesn't work well
```


## lg lm
```{r}
model2<- lm(formula = log(real.salary,10) ~ Age  + Total.rebound.percetage + 
    Assist.percetage + Steal.percetage + Field.goal.percetage + Free.throw.percetage + height + Average.points, 
data = Season_salary)

summary(model2)
plot(model2)


summary(lglm)
plot(lglm)


x <- data.frame(Season_salary$Age,Season_salary$Total.rebound.percetage ,
    Season_salary$Assist.percetage , Season_salary$Steal.percetage , Season_salary$Field.goal.percetage , 
    Season_salary$Free.throw.percetage , Season_salary$height, Season_salary$Average.points)
X <- data.matrix(x)



set.seed(123)
training.samples <- real.salary %>%
  createDataPartition(p = 0.7, list = FALSE)
train.data  <- x[training.samples, ]
train.y <- (real.salary[training.samples])
test.y <- real.salary[-training.samples]
test.data <- x[-training.samples, ]
# Build the model
model2 <- lm(log(real.salary,10) ~ X, data = train.data)
# Make predictions and compute the R2, RMSE and MAE
predictions <- model2 %>% predict(test.data)

#MSE:
#data.frame( R2 = R2(predictions, real.salary),
model2_RMSE = RMSE(predictions, real.salary)
model2_MSE = mse(predictions, real.salary)



k10_cv = trainControl(method="cv", number=10)
ols_model2 = train(real.salary ~ Age + Player.efficiency.rating + Total.rebound.percetage + 
    Assist.percetage + Steal.percetage + Field.goal.percetage + 
    Free.throw.percetage + height + weight + Average.points, 
    data = Season_salary,trControl = k10_cv, method = "lm")
ols_model2$results
```


## Ridge & Lasso
```{r}
x <- data.frame(Season_salary$Age,Season_salary$Total.rebound.percetage ,
    Season_salary$Assist.percetage , Season_salary$Steal.percetage , Season_salary$Field.goal.percetage , 
    Season_salary$Free.throw.percetage , Season_salary$height, Season_salary$Average.points)
X <- data.matrix(x)


#model 4 salary model ridge & lasso

Ridge <- cv.glmnet(X,log(real.salary,10), alpha=0)
print(Ridge$lambda.min)

Lasso <- cv.glmnet(X,log(real.salary,10), alpha=1)
print(Lasso$lambda.min)

#### Plot of the coefficient values

plot(Ridge)
plot(Lasso)



#### Using the “best” lambda value

BestLasso <- glmnet(X,log(real.salary,10),alpha = 1, lambda = Lasso$lambda.min)
BestLasso$beta


BestRidge <- glmnet(X,log(real.salary,10),alpha = 0, lambda = Ridge$lambda.min)
BestRidge$beta

ridge_model = train(real.salary ~ Age  + Total.rebound.percetage + 
    Assist.percetage + Steal.percetage + Field.goal.percetage + 
    Free.throw.percetage + height + Average.points, 
    data = Season_salary,trControl = k10_cv, method = "ridge")
ridge_model$results

lasso_model = train(real.salary ~ Age  + Total.rebound.percetage + 
    Assist.percetage + Steal.percetage + Field.goal.percetage + 
    Free.throw.percetage + height  + Average.points, 
    data = Season_salary,trControl = k10_cv, method = "lasso")
lasso_model$results

```

## Principle Component Model
```{r}
x <- data.frame(Season_salary$Age,Season_salary$Player.efficiency.rating , Season_salary$Total.rebound.percetage ,
    Season_salary$Assist.percetage , Season_salary$Steal.percetage , Season_salary$Field.goal.percetage , 
    Season_salary$Free.throw.percetage , Season_salary$height , Season_salary$weight , Season_salary$Average.points)


pca.x <- data.frame(Season_salary$Age,Season_salary$Player.efficiency.rating , Season_salary$Total.rebound.percetage ,
    Season_salary$Assist.percetage , Season_salary$Steal.percetage , Season_salary$Field.goal.percetage , 
    Season_salary$Free.throw.percetage , Season_salary$height , Season_salary$weight , Season_salary$Average.points)


pca.full <- data.frame(real.salary,pca.x)

pca.X <- data.matrix(pca.x)
pca.FULL <- data.matrix(pca.full)

fa.parallel(pca.x, fa = 'pc', n.iter = 100, show.legend = FALSE) #Scree plot

# pca = PCA(pca.x, scale.unit=TRUE, ncp=3, graph=TRUE)
# pca = PCA(pca.x, scale.unit=FALSE, ncp=3, graph=TRUE)
# dimdesc(pca, axes=c(1,2))
da.panelellipses=panellipse(pca.full, col.p=1, col.j=2, firstvar=3)

pca.method1 <- principal(pca.x, nfactors = 3, rotate = "none")
pca.method1

pca.model <- lm(real.salary ~ pca.method1$scores[,1] + pca.method1$scores[,2] + pca.method1$scores[,3])
summary(pca.model)

pca.cv <- data.frame(real.salary, pca.method1$scores[,1], pca.method1$scores[,2], pca.method1$scores[,3])
names(pca.cv) <- c("real.salary", "pc1", "pc2", "pc3")

pca.result = train(real.salary ~ pc1 + pc2, data= pca.cv, trControl = k10_cv, method = "lm")
pca.result$results



# Draw conclusion: which model is the best one
```

#### Nonlinear - KNN
```{r}
?knnreg
?plot


fit <- knnreg(train.data, train.y, k = 5)

plot(test.y, predict(fit, test.x))



line(test.y==predict(fit, test.data))

knn = train(real.salary ~ Age + Player.efficiency.rating + Total.rebound.percetage + 
    Assist.percetage + Steal.percetage + Field.goal.percetage + 
    Free.throw.percetage + height + weight + Average.points, 
    data = Season_salary,trControl = k10_cv, method = "knn")
knn$results
# Draw conclusion: which model is the best one
```


## Part 3: Hall of Fame Prediction:
```{r}
# Hall of Fame Prediction
Seasondata_After1980$Player <- as.character(Seasondata_After1980$Player)
a <- Seasondata_After1980
a$Player = as.character(gsub("[\\*,]", "", a$Player))

a$hof <- as.character(a$Player!=Seasondata_After1980$Player)

#====================================================
# Build HoF Training Sample:
hof_players <- subset(a, a$hof=="TRUE")
regular_players <- subset(a, a$hof=="FALSE")
regular_players <- subset(regular_players, regular_players$Year < 1995)


hof_players$hof.poss <- 1
hof_players <- hof_players[,-c(52,53)]
hof_players <- hof_players[,-c(1,2,4,6,8)]
hof_players <- hof_players[,-c(2)]
hof_players <- hof_players[,-c(2,3)]
hof_players <- as.data.frame(hof_players)

regular_players$hof.poss <- 0
regular_players <- regular_players[,-c(52,53)]
regular_players <- regular_players[,-c(1,2,4,6,8)]
regular_players <- regular_players[,-c(2)]
regular_players <- regular_players[,-c(2,3)]
regular_players <- as.data.frame(regular_players)


# hof_players <- arrange(hof_players, hof_players$Player)
hof_players<- mtable(hof_players, bycol = "Player", firstvarcol = 2)
hof_players <- as.data.frame(hof_players)
hof_players <- subset(hof_players, hof_players$Points > 100)
hof_players$hof.poss <- 1
hof_players <- as.data.frame(hof_players)
hof_players <- hof_players[complete.cases(hof_players),]



regular_players<- mtable(regular_players, bycol = "Player", firstvarcol = 2)
regular_players <- as.data.frame(regular_players)
#regular_players <- subset(regular_players, regular_players$Points > 100)
regular_players$hof.poss <- 0
regular_players <- as.data.frame(regular_players)
regular_players <- regular_players[complete.cases(regular_players),]

sample <- rbind(hof_players, regular_players)
sample.x <- sample[,-43]
#====================================================
# Build Prediction Sample:
regular_players1 <- subset(a, a$hof=="FALSE")
regular_players1 <- subset(regular_players1, regular_players1$Year >= 1995)

regular_players1 <- regular_players1[,-c(1,2,4,5,6,7,8,52)]
regular_players1 <- regular_players1[,-2]
regular_players1 <- mtable(regular_players1, bycol = "Player", firstvarcol = 2)
regular_players1 <- as.data.frame(regular_players1)

regular_players1 <- regular_players1[,-43]
regular_players1 <- regular_players1[complete.cases(regular_players1),]
prediction.x <- regular_players1


#====================================================

fa.parallel(sample.x, fa = 'pc', n.iter = 100, show.legend = FALSE)
pca = PCA(sample.x, scale.unit=TRUE, ncp=6, graph=TRUE)


hof.method <- principal(sample.x, nfactors = 6, rotate = "none")
hof.method


logit.dataset <- data.frame(sample$hof.poss, hof.method$scores[,1], hof.method$scores[,2], hof.method$scores[,3], hof.method$scores[,4], hof.method$scores[,5], hof.method$scores[,6])
names(logit.dataset) <- c("hof.poss", "pc1", "pc2","pc3","pc4","pc5","pc6")

hof.model <- glm(hof.poss~ pc1 + pc2 + pc3 + pc4 + pc5 + pc6 , data = logit.dataset, family = binomial)
summary(hof.model)  # Logistic Model



#fa.parallel(prediction.x, fa = 'pc', n.iter = 100, show.legend = FALSE)
#pca = PCA(prediction.x, scale.unit=TRUE, ncp=6, graph=TRUE)

predict.method <- principal(prediction.x, nfactors = 6, rotate = "none")
predict.method

prediction.x <- data.frame(predict.method$scores[,1], predict.method$scores[,2], predict.method$scores[,3], predict.method$scores[,4], predict.method$scores[,5], predict.method$scores[,6])
names(prediction.x) <- c("pc1", "pc2","pc3","pc4", "pc5", "pc6")

# logit.model <- glm(hof.poss~., data = hof_players, family=binomial(link="logit"))



prediction.x$hof.poss <- plogis(predict(hof.model, prediction.x))

prediction.x



hof_prediction <- setDT(prediction.x, keep.rownames = TRUE)[]
hof_prediction <- hof_prediction[,-c(2,3,4,5,6,7)]
hof_prediction <- hof_prediction[order(hof_prediction$hof.poss),]
hof_prediction <- hof_prediction[nrow(hof_prediction):1]
hof_prediction[1:20,]
```


